# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
* Dataset used for this Assessment (activity.zip) is present in the forked git repository.
* Lets load the csv file if already uncompressed, else lets unzip the "activity.zip" and load the dataset.

```{r Loading_Data, echo = TRUE}
fileName <- "activity.csv"
if(file.exists(fileName)){
    data <- read.csv(fileName)
}else{
    unzip("activity.zip")
    data <- read.csv(fileName)
}
```
* The date column is converted to Date type
* The inverval column is converted to factor type. 

```{r Preprocessing, echo = TRUE}
data$date <- as.Date(data$date) #Converting data column to Date type
data$interval <- as.factor(data$interval) #Converting interval column to factor type
```

* Lets look at the structure and summary of data
```{r Look_at_the_Data, echo = TRUE}
str(data)
summary(data)
```

## What is mean total number of steps taken per day?

* calculating total number of steps taken per day
```{r Aggregation, echo = TRUE}
steps_taken_per_day <- aggregate(steps ~ date, data, sum)
```

* Lets see the histogram of steps taken per day
```{r Histogram, fig.width = 6, fig.height= 6, echo = TRUE}
hist(steps_taken_per_day$steps, col = "green")
```

* **Mean** of steps taken per day
```{r Mean, echo = TRUE}
mean(steps_taken_per_day$steps)
```

* **Median** of steps taken per day
```{r Median, echo = TRUE}
median(steps_taken_per_day$steps)
```

## What is the average daily activity pattern?
* Calculating average number of steps taken per interval
```{r stpes_per_interval, echo = TRUE}
steps_per_interval <- aggregate(steps ~ interval, data, mean)
```

* Plotting timeseries of average daily activity pattern
```{r timeseries, fig.width = 6, fig.height=6}
plot(steps_per_interval$steps, col = "green", type = "l", main = "Timeseries of Daily Pattern", ylab = "Average stpes per interval")
```

* Maximum number of steps in which 5 minute interval?

```{r}
steps_per_interval[max(steps_per_interval$steps),1]
```

## Imputing missing values

Imputing a non-sophasticated method for NAs, Replacing NA's with zero

```{r imputing_missing, echo = TRUE}
steps <- data$steps
steps[is.na(steps)] <- 0
data$steps <- steps
```

* calculating total number of steps taken per day
```{r new_Aggregation, echo = TRUE}
steps_taken <- aggregate(steps ~ date, data, sum)
```

* Lets see the new histogram of steps taken per day
```{r new_Histogram, fig.width = 6, fig.height= 6, echo = TRUE}
hist(steps_taken$steps, col = "green")
```

* **Mean** of steps taken per day
```{r new_Mean, echo = TRUE}
mean(steps_taken$steps)
```

* **Median** of steps taken per day
```{r new_Median, echo = TRUE}
median(steps_taken$steps)
```

## Are there differences in activity patterns between weekdays and weekends?

* Creating column weekday
```{r}
data$weekday <- as.factor(weekdays(data$date))
```

 * Creating one more column for weekday and weekend
```{r}
data$day <- NA
data[data$weekday %in% c("Saturday", "Sunday"),5] <- "Weekend"
data[!data$weekday %in% c("Saturday", "Sunday"),5] <- "Weekday"
data$day <- as.factor(data$day)
```

* Comparison of activities of weekday and weekend
```{r comparison_plot, fig.height=6, fig.width=6, echo = TRUE}
library(lattice)
panel.smoother <- function(x, y) {
   panel.loess(x, y)  # show smoothed line 
}
data$interval <- as.numeric(data$interval)
xyplot(steps ~ interval |day, data = data, layout = c(1,2), type = "l", ylab = "Number of steps")

```
